timesliced_random_cell_selection

### 2018-11-19 src/env_data_prep.Rmd
background bathy data

```{r}
data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE)
data_aea$bottom_depth <- data_aea$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth. This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
head(data_aea)
```

>>>>

```{r}
data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE) #5 = decimalLongitude, 4 = decimalLatitude
data_aea$bottom_depth <- data_aea$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth. This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
head(data_aea)
```

### 2018-11-19 src/timesliced_random_cell_selection.Rmd
background bathy data

```{r pre-install & load packages, include=FALSE}
install.packages("pacman")
library(pacman)
p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap)
```

@@

>>>>

```{r pre-install & load packages, include=FALSE}
install.packages("pacman")
library(pacman)
p_load(raster, dplyr, data.table, sp, adehabitatHR, rgdal, rworldmap, ncdf4, marmap)
```

@@

the depth layers in the netCDF are literal slices. Currently I have listed the depthlayerno (which is literally which layer in the netcdf). I should also add the depth_layer name which gives the depth the slice is supposed to represent. Data is stored in depth_bin.csv

load the depth bin csv
```{r}
depthbin <- read.csv("../data/env/depth_bin.csv", header = TRUE) 
head(depthbin)
```

so basically what i want to do is, where depthbin$layer_no = background_all$depthlayerno, extract depthbin$layer_name and add to background_all$depth_layer. Can use the merge function.
First need to change depthbin col names to match background_all

```{r}
names(depthbin)[names(depthbin)=="layer_no"] <- "depthlayerno"
names(depthbin)[names(depthbin)=="layer_name"] <- "depth_layer" #this one is just so i don't have to change the column name in the background_all file later
head(depthbin)
```
```{r}
background_all <- merge(x = background_all, y = depthbin[ , c("depthlayerno", "depth_layer")], by.x = ("depthlayerno"), all.x = TRUE)
head(background_all)
```

just check some random rows...

```{r}
randcheck <- background_all[sample(nrow(background_all), 10), ]
randcheck
```
all good - save to .csv

```{r}
write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname.csv", row.names = FALSE)
```

# Bathymetric

bottom depth from [GEBCO](www.gebco.net)

```{r}
bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
summary(bat)
```

create pretty custom color palette and then plot the bathy laye
```{r}
blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
plot(bat, n = 1, image = TRUE, bpal = blues(100))
```
lovely jubbely

add the species data...
```{r}
plot(bat, n = 1, image = TRUE, bpal = blues(100))
points(background_all$decimalLongitude, background_all$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
```
and now get depth... note all the points and bathy data is wgs!

```{r}
background_all$bottom_depth <- get.depth(bat, background_all[ ,10:11], locator = FALSE)
background_all$bottom_depth <- background_all$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth. This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao_depthname_bath.csv", row.names = FALSE)
head(background_all)
```

# observations by cell (in total, and also by time step)

as the title says... how many observations happeend in each cell:
- in total
- in each yyyymm

```{r observation by cell - total}
library(plyr) #to use the count function
obs_by_cell <- count(data_aea, "cell_id")
obs_by_cell
write.csv(obs_by_cell, file = "../output/bio/no_observations_cell.csv")
```

```{r}
obs_cell_yymm <- count(data_aea, c("cell_id", "year", "month"))
obs_cell_yymm
write.csv(obs_cell_yymm, file = "../output/bio/no_observations_cell_yymm.csv")
```
Now attach these values to data_aea....

total obs for the whole time period
```{r}
data_aea <- merge(data_aea, obs_by_cell, by="cell_id")
head(data_aea)
```
rename freq column to something meaningful

```{r}
names(data_aea)[names(data_aea)=="freq"] <- "total_cell_obs"
head(data_aea)
```

time-sliced
```{r}
data_aea <- merge(data_aea, obs_cell_yymm, by.x=c("cell_id", "year", "month"), by.y=c("cell_id", "year", "month"))
names(data_aea)[names(data_aea)=="freq"] <- "yymm_cell_obs"
head(data_aea)
```

write to .csv
```{r}
write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno.csv", row.names = FALSE)
```
`
